### 2020-09-22 LaTeX/sections/paper.tex
More edits

\paragraph{Contributions.} We design and implement \cm, a fully on-chain call market in Solidity---a high-level programming language for Ethereum. We choose Ethereum as a `hostile' environment for an orderbook: Ethereum is public (strongest adversary) and slow (lower-bound benchmark) relative to more scalable blockchains, including the anticipated Ethereum 2.0. \cm is intended as a simple module that developers can modify as they choose. By adding \cm to an ERC20 (or similar) token contract, everything that is needed for buying/selling tokens is immediately available on chain without involving any third party exchanges. \cm is non-custodial, transparent, and mitigates front-running attacks that are generally possible on-chain.

@@

\paragraph{Execution systems.} There are three main approaches to arranging a trade~\cite{Har03}. In a \emph{quote-driven} market, a dealer uses their own inventory to offer a price for buying or selling an asset. In a \emph{brokered exchange}, a broker finds a buyer and seller. In an \emph{order-driven} market, offers to buy (\emph{bids}) and sell (\emph{offers}/\emph{asks}) from many traders are placed as orders in an orderbook. Order-driven markets can be \emph{continuous}, with buyers/sellers at any time adding orders to the orderbook (\emph{makers}) or executing against an existing order (\emph{takers}); or they can be \emph{called}, where all traders submit orders within a window of time and orders are matched in a batch like an auction.

@@

\paragraph{Central Exchanges (CEX).} Traditional financial markets (\eg NYSE and NASDAQ) use order-matching systems to arrange trades. An exchange will list one or more assets (stocks, bonds, derivatives, or more exotic securities) to be traded with each given its own orderbook priced in a currency (\eg USD). Exchanges for blockchain-based assets (also called crypto assets by enthusiasts) can operate the same way, using a centralized exchange (CEX) design where a firm operates the platform as a trusted third party in every aspect: custodianship over assets/currency being traded, exchanging assets fairly, and offering the best possible price execution. Security breaches and fraud (\eg MtGox~\cite{TheHisto45:online}, QuadrigaCX~\cite{SEBIOrde83:online}, and many others) in centralized exchanges have become a common source of lost funds for users, while accusations of unfair trade execution have been levelled but are difficult to prove.

\paragraph{On-chain Orderbooks.} For trades between two blockchain-based assets (\eg a digital asset priced with a cryptocurrency, stablecoin, or second digital asset), order matching can be performed `on-chain' by deploying the order-matching system either on a dedicated blockchain or inside a decentralized application (DApp; \textit{a.k.a.} smart contract). In this model, traders entrust their assets to an autonomously operating DApp with known source code instead of a third party custodian that can abscond with or lose the funds. The trading rules will operate as coded, clearing and settling can be guaranteed, and order submission is handled by the blockchain---a reasonably fair and transparent system (but see front-running below). Finally, anyone can create an on-chain orderbook for any asset (on the same chain) at any time. While they sound ideal, performance is a substantial issue and the main subject of this paper.

In this paper, we focus on benchmarking an orderbook for a public blockchain (\eg Ethereum). Ethereum is widely-used and we stand to learn the most from working in a performance-hostile environment (\ie Ethereum is good lower-bound). Exchanges could be given their own dedicated blockchain, where trade execution logic can be moved at the consensus level. Permissioned blockchains (\eg NASDAQ Linq, tZero) can also increase execution time and throughput, possibly with some reduction in transparency and trust.

\paragraph{On-chain Dealers.} Another on-chain advantage is that other smart contracts, not just human users, can initiate trades, enabling broader decentralized finance (DeFi) applications. This has fuelled a resurgence in on-chain exchange but through a quote-driven design rather than an order-driven one. Automated market makers  (\eg Uniswap) have all the trust advantages of an on-chain orderbook, plus they are very efficient relative to an on-chain orderbook. The trade-off is that they operate as a dealer---the DApp exchanges assets from its own inventory. This inventory is loaded into the DApp by an investor who will not make money on the trades themselves, but hope for long term profit through trading fees. By contrast, an orderbook requires no upfront inventory and does not have to charge trading fees (but can). Finally, there is a complicated difference in their price dynamics (\eg market impact of a trade, slippage between the best bid/ask and actual average execution price, \etc)---deserving of an entire research paper to precisely define. We leave it as an assertion that with equal liquidity, orderbooks have more favourable price dynamics for traders.

\paragraph{Hybrid Designs.} Before on-chain dealers became prominent in the late 2010s, the most popular design was hybrid order-driven exchanges with some trusted off-chain components and some on-chain functionality. Some (\eg EtherDelta) were envisioned as operating fully on-chain, but performance limitations drove developers to move key components, such as the order matching system, off-chain to a centralized database. A landscape of DEX designs exist: many avoiding taking custodianship of assets off-chain, and virtually all (for order-driven markets) operate the orderbook itself off-chain. A non-custodial DEX solves the big issue of a CEX---the operator stealing the funds---however trade execution is still not provably fair. A mitigation is to issue a proof of correct execution to the DApp (\eg Loopring) but these proofs have blindspots (discussed in section~\ref{sec:rollups}). \cm offers key advantages in this model.

\subsection{Related Work}

\paragraph{Blockchain Limitations and Solution.} While an orderbook is a ledger and blockchains provide a distributed ledger, it is not straightforward to drop a continuous-time orderbook onto a blockchain. An older 2014 paper~\cite{clark2014decentralizing} on the `Princeton prediction market'~\cite{Bra13} motivates our work. The authors observe the following limitations of on-chain continous orderbooks: block intervals are slow and not continuous, there is no support for accurate time-stamping, transactions can be dropped or reordered by miners, fast traders can react to submitted orders/cancelations when broadcast to network but not in a block and have their orders appear first (as examined in later work on front-running: \cite{eskandari2019sok,daian2019flash}).

\paragraph{Call Markets.} The researchers propose using a call market instead of a continuous-time market~\cite{clark2014decentralizing}. Orders are collected and placed into the orderbook over a window of time (\eg 1 or more blocks), then the market is closed and the orders are processed in batch: the best bids are matched to the best asks in order. If the prices overlap, the miner keeps the difference (which they could extract anyways through front-running). Call markets largely side-step front-running attacks from other traders because reordering trades has no impact (discussed more in section~\ref{sec:front}). The paper does not include an implementation and was envisioned as an alt-coin (Ethereum was still in development in 2014) with market closing being done by the miners themselves as part of the blockchain logic.

Large exchanges, like the NYSE and NASDAQ, use a two minute call market every day at opening and closing time (in between, the exchange runs as a continuous time market). Other exchanges, called crossing networks, also operate as a call market at various times throughout the trading day.\footnote{A crossing network uses a secondary market for determining the closing price. Many prominent crossing networks are operated internally within a brokerage for its clients, and often as a `	dark pool' with an unpublished orderbook.} Call market are studied widely in finance~\cite{Har03}. Time-sensitive traders submit orders early, especially in crossing networks that close at a randomly determined time (traders risk missing the call if they wait too long). A blockchain happens to provide this function naturally, as blocks are published unpredictably. Price-sensitive traders wait to base their pricing off the already submitted orders and do not mind missing a call if it obtains them a better price.

\paragraph{Other Academic Literature.} There are numerous industry projects on blockchain-based exchanges and orderbooks but most of the academic literature is on topics that are related but tangential to the mechanics of trade execution. Early literature considered trade execution under encryption (\ie dark markets)~\cite{TP07,YSLT10,TW12}. A number of projects consider structuring futures and derivatives in smart contracts---Velocity ~\cite{eskandari2017feasibility}, Findel ~\cite{biryukov2017findel}, FuturesMex~\cite{massacci2018futuresmex}---but once issued, the derivative can be traded using exchanges. Atomic swaps (\ie payment vs delivery) are necessary for settling trades and some general approaches include Arwen~\cite{heilman2020arwen} and Tesseract~\cite{bentov2017tesseract}.

@@

\subsubsection{Enqueue performance.} We implemented, deployed, and tested each priority queue using Truffle and Ganache.\footnote{Github: Link removed for anonymity.} We tried a variety of tests (including testing the full call market with each varient) with consistent results in performance. A simple test to showcase the performance profile is shown in Figure~\ref{fig:random_insertion}. We simply enqueue 50 integers chosen at random from a fixed interval in each PQ variant. The bigger the PQ gets, the longer enqueue takes---a linear increase for the linked list variants, and logarithmic for the heap variants.

@@

\subsubsection{Dequeue Performance.} For each PQ variant storing 50 random integers, the \texttt{Dequeue()} function is iterated until the data structure is empty. The total gas cost for fully dequeuing the priority queue variants is outlined in Table~\ref{tab:PQUnitTests}. These tests are performed using the following Ethereum gas metrics: block gas limit $=$11,741,495 and 1 gas $=$ 56 gwei.\footnote{EthStats (July 2020): \url{https://ethstats.net/}} Dequeuing removes data from the contract's storage. Recall this results in a gas refund. Based on our manual estimates (EVM does not expose the refund counter), every variant receives the maximum gas refund possible (\ie half the total cost of the transaction). In other words, each of them actually consumes twice the \texttt{gasUsed} amount in gas before the refund, however none of them are better or worse based on how much of a refund they generate.

@@

One consequence of a linked list is that a new contract is created for every node in the list. Beyond being expensive for adding new nodes (a cost that will be bared by the trader in a call market), it also leaves a large footprint on the Ethereum blockchain, especially if we leave the nodes on the blockchain in perpetuity (\ie we just update the head node of the list and leave the previous head `dangling.'). However in a PQ, nodes are only removed from the head of the list; thus the node contracts could be `destroyed' one by one using an extra operation, \texttt{SELFDESTRUCT}, in the \texttt{dequeue()} function. As shown in Table~\ref{tab:cleaning}, the refund from doing this outweighs to the cost of the extra computation: gas costs are reduced from 721K to 557K.  This suggests a general principle: cleaning up after yourself will pay for itself in gas refunds. Unfortunately, this is not universally true as shown by applying the same principle to the linked list with mapping.

@@

\paragraph{Front-running.} The primary motivation for using a call market, as opposed to a continuous time orderbook, is to mitigate front-running attacks~\cite{clark2014decentralizing,eskandari2019sok,daian2019flash}. Consider a sequence of 100 orders submitted within a window of time to market A, and the same sequence randomly shuffled and submitted to market B. If A and B are continuous, the orders that are executed in A could be quite different from B. In a call market, the outcome of A and B will be exactly equivalent. For this reason, there is no threat from miners reordering transactions or users offering higher gas rates to have their orders executed before other orders already broadcasted.

>>>>

\paragraph{Contributions.} We design and implement \cm, a fully on-chain call market in Solidity---a high-level programming language for Ethereum. We choose Ethereum as a `hostile' environment for an order book: Ethereum is public (strongest adversary) and slow (lower-bound benchmark) relative to more scalable blockchains, including the anticipated Ethereum 2.0. \cm is intended as a simple module that developers can modify as they choose. By adding \cm to an ERC20 (or similar) token contract, everything that is needed for buying/selling tokens is immediately available on chain without involving any third party exchanges. \cm is non-custodial, transparent, and mitigates front-running attacks that are generally possible on-chain.

@@

\paragraph{Execution systems.} There are three main approaches to arranging a trade~\cite{Har03}. In a \emph{quote-driven} market, a dealer uses their own inventory to offer a price for buying or selling an asset. In a \emph{brokered exchange}, a broker finds a buyer and seller. In an \emph{order-driven} market, offers to buy (\emph{bids}) and sell (\emph{offers}/\emph{asks}) from many traders are placed as orders in an order book. Order-driven markets can be \emph{continuous}, with buyers/sellers at any time adding orders to the order book (\emph{makers}) or executing against an existing order (\emph{takers}); or they can be \emph{called}, where all traders submit orders within a window of time and orders are matched in a batch like an auction.

@@

\paragraph{Central Exchanges (CEX).} Traditional financial markets (\eg NYSE and NASDAQ) use order-matching systems to arrange trades. An exchange will list one or more assets (stocks, bonds, derivatives, or more exotic securities) to be traded with each given its own order book priced in a currency (\eg USD). Exchanges for blockchain-based assets (also called crypto assets by enthusiasts) can operate the same way, using a centralized exchange (CEX) design where a firm operates the platform as a trusted third party in every aspect: custodianship over assets/currency being traded, exchanging assets fairly, and offering the best possible price execution. Security breaches and fraud (\eg MtGox~\cite{TheHisto45:online}, QuadrigaCX~\cite{SEBIOrde83:online}, and many others) in centralized exchanges have become a common source of lost funds for users, while accusations of unfair trade execution have been levelled but are difficult to prove.

\paragraph{On-chain Order Books.} For trades between two blockchain-based assets (\eg a digital asset priced with a cryptocurrency, stablecoin, or second digital asset), order matching can be performed `on-chain' by deploying the order-matching system either on a dedicated blockchain or inside a decentralized application (DApp; \textit{a.k.a.} smart contract). In this model, traders entrust their assets to an autonomously operating DApp with known source code instead of a third party custodian that can abscond with or lose the funds. The trading rules will operate as coded, clearing and settling can be guaranteed, and order submission is handled by the blockchain---a reasonably fair and transparent system (but see front-running below). Finally, anyone can create an on-chain order book for any asset (on the same chain) at any time. While they sound ideal, performance is a substantial issue and the main subject of this paper.

In this paper, we focus on benchmarking an order book for a public blockchain (\eg Ethereum). Ethereum is widely-used and we stand to learn the most from working in a performance-hostile environment (\ie Ethereum is good lower-bound). Exchanges could be given their own dedicated blockchain, where trade execution logic can be moved at the consensus level. Permissioned blockchains (\eg NASDAQ Linq, tZero) can also increase execution time and throughput, possibly with some reduction in transparency and trust.

\paragraph{On-chain Dealers.} Another on-chain advantage is that other smart contracts, not just human users, can initiate trades, enabling broader decentralized finance (DeFi) applications. This has fuelled a resurgence in on-chain exchange but through a quote-driven design rather than an order-driven one. Automated market makers  (\eg Uniswap) have all the trust advantages of an on-chain order book, plus they are very efficient relative to an on-chain order book. The trade-off is that they operate as a dealer---the DApp exchanges assets from its own inventory. This inventory is loaded into the DApp by an investor who will not make money on the trades themselves, but hope for long term profit through trading fees. By contrast, an order book requires no upfront inventory and does not have to charge trading fees (but can). Finally, there is a complicated difference in their price dynamics (\eg market impact of a trade, slippage between the best bid/ask and actual average execution price, \etc)---deserving of an entire research paper to precisely define. We leave it as an assertion that with equal liquidity, order books have more favourable price dynamics for traders.

\paragraph{Hybrid Designs.} Before on-chain dealers became prominent in the late 2010s, the most popular design was hybrid order-driven exchanges with some trusted off-chain components and some on-chain functionality. Some (\eg EtherDelta) were envisioned as operating fully on-chain, but performance limitations drove developers to move key components, such as the order matching system, off-chain to a centralized database. A landscape of DEX designs exist: many avoiding taking custodianship of assets off-chain, and virtually all (for order-driven markets) operate the order book itself off-chain. A non-custodial DEX solves the big issue of a CEX---the operator stealing the funds---however trade execution is still not provably fair. A mitigation is to issue a proof of correct execution to the DApp (\eg Loopring) but these proofs have blindspots (discussed in section~\ref{sec:rollups}). \cm offers key advantages in this model.

\subsection{Related Works}

\paragraph{Blockchain Limitations and Solution.} While an order book is a ledger and blockchains provide a distributed ledger, it is not straightforward to drop a continuous-time order book onto a blockchain. An older 2014 paper~\cite{clark2014decentralizing} on the `Princeton prediction market'~\cite{Bra13} motivates our work. The authors observe the following limitations of on-chain continuous order books: block intervals are slow and not continuous, there is no support for accurate time-stamping, transactions can be dropped or reordered by miners, fast traders can react to submitted orders/cancelations when broadcast to network but not in a block and have their orders appear first (as examined in later work on front-running: \cite{eskandari2019sok,daian2019flash}).

\paragraph{Call Markets.} The researchers propose using a call market instead of a continuous-time market~\cite{clark2014decentralizing}. Orders are collected and placed into the order book over a window of time (\eg 1 or more blocks), then the market is closed and the orders are processed in batch: the best bids are matched to the best asks in order. If the prices overlap, the miner keeps the difference (which they could extract anyways through front-running). Call markets largely side-step front-running attacks from other traders because reordering trades has no impact (discussed more in section~\ref{sec:front}). The paper does not include an implementation and was envisioned as an alt-coin (Ethereum was still in development in 2014) with market closing being done by the miners themselves as part of the blockchain logic.

Large exchanges, like the NYSE and NASDAQ, use a two minute call market every day at opening and closing time (in between, the exchange runs as a continuous time market). Other exchanges, called crossing networks, also operate as a call market at various times throughout the trading day.\footnote{A crossing network uses a secondary market for determining the closing price. Many prominent crossing networks are operated internally within a brokerage for its clients, and often as a `	dark pool' with an unpublished order book.} Call market are studied widely in finance~\cite{Har03}. Time-sensitive traders submit orders early, especially in crossing networks that close at a randomly determined time (traders risk missing the call if they wait too long). A blockchain happens to provide this function naturally, as blocks are published unpredictably. Price-sensitive traders wait to base their pricing off the already submitted orders and do not mind missing a call if it obtains them a better price.

\paragraph{Other Academic Literature.} There are numerous industry projects on blockchain-based exchanges and order books but most of the academic literature is on topics that are related but tangential to the mechanics of trade execution. Early literature considered trade execution under encryption (\ie dark markets)~\cite{TP07,YSLT10,TW12}. A number of projects consider structuring futures and derivatives in smart contracts---Velocity ~\cite{eskandari2017feasibility}, Findel ~\cite{biryukov2017findel}, FuturesMex~\cite{massacci2018futuresmex}---but once issued, the derivative can be traded using exchanges. Atomic swaps (\ie payment vs delivery) are necessary for settling trades and some general approaches include Arwen~\cite{heilman2020arwen} and Tesseract~\cite{bentov2017tesseract}.

@@

\subsubsection{Enqueue performance.} We implemented, deployed, and tested each priority queue using Truffle and Ganache.\footnote{Github: Link removed for anonymity.} We tried a variety of tests (including testing the full call market with each variant) with consistent results in performance. A simple test to showcase the performance profile is shown in Figure~\ref{fig:random_insertion}. We simply enqueue 50 integers chosen at random from a fixed interval in each PQ variant. The bigger the PQ gets, the longer enqueue takes---a linear increase for the linked list variants, and logarithmic for the heap variants.

@@

\subsubsection{Dequeue Performance.} For each PQ variant storing 50 random integers, the \texttt{Dequeue()} function is iterated until the data structure is empty. The total gas cost for fully dequeuing the priority queue variants is outlined in Table~\ref{tab:PQUnitTests}. These tests are performed using the following Ethereum gas metrics: block gas limit $=$11,741,495 and 1 gas $=$ 56 Gwei.\footnote{EthStats (July 2020): \url{https://ethstats.net/}} Dequeuing removes data from the contract's storage. Recall this results in a gas refund. Based on our manual estimates (EVM does not expose the refund counter), every variant receives the maximum gas refund possible (\ie half the total cost of the transaction). In other words, each of them actually consumes twice the \texttt{gasUsed} amount in gas before the refund, however none of them are better or worse based on how much of a refund they generate.

@@

One consequence of a linked list is that a new contract is created for every node in the list. Beyond being expensive for adding new nodes (a cost that will be bared by the trader in a call market), it also leaves a large footprint on the Ethereum blockchain, especially if we leave the nodes on the blockchain in perpetuity (\ie we just update the head node of the list and leave the previous head `dangling.'). However in a PQ, nodes are only removed from the head of the list; thus the node contracts could be `destroyed' one by one using an extra operation, \texttt{SELFDESTRUCT}, in the \texttt{Dequeue()} function. As shown in Table~\ref{tab:cleaning}, the refund from doing this outweighs to the cost of the extra computation: gas costs are reduced from 721K to 557K.  This suggests a general principle: cleaning up after yourself will pay for itself in gas refunds. Unfortunately, this is not universally true as shown by applying the same principle to the linked list with mapping.

@@

\paragraph{Front-running.} The primary motivation for using a call market, as opposed to a continuous time order book, is to mitigate front-running attacks~\cite{clark2014decentralizing,eskandari2019sok,daian2019flash}. Consider a sequence of 100 orders submitted within a window of time to market A, and the same sequence randomly shuffled and submitted to market B. If A and B are continuous, the orders that are executed in A could be quite different from B. In a call market, the outcome of A and B will be exactly equivalent. For this reason, there is no threat from miners reordering transactions or users offering higher gas rates to have their orders executed before other orders already broadcasted.
